Informing IPCC accounting of forest carbon using the global forest carbon database (ForC v4.0)

### 2022-04-16 doc/manuscript/ESSDmanuscript.Rmd
appendix arrangement

bibliography: references.bib
running:
  title: IPCC forest C accounting with ForC
  author: Williams et al.
competinginterests: |
  The authors declare no competing interests.
# OPTIONAL:
algorithms: true
# See https://publications.copernicus.org/for_authors/licence_and_copyright.html, normally used for transferring the copyright, if needed. 
# Note: additional copyright statements for affiliated software or data need to be placed in the data availability section. 
copyrightstatement: |
### The author's copyright for this publication is transferred to institution/company. 
### Note: unless stated otherwise, software and data affiliated with the manuscript are assumed to be published under the same licence as the article (currently Creative Commons 4.0)
availability:
  #  use this to add a statement when having only software code available
  #  use this to add a statement when having only data sets available
  #code: |
  #data: |
  codedata: |
    use this to add a statement when having data sets and software code available
authorcontribution: |
  (fill this in)
disclaimer: |
acknowledgements: |
  Thank you to all researchers who collected and published the data contained in ForC, and to all research assistants and collaborators who have helped to build the database.
  Funding for this study was provided by Bezos Earth Fund to The Nature Conservancy, the Institute for Global Environmental Strategies, WLS(?)
appendix: |
   \section{Mapping ForC to EFDB}

   CURRENT TABLE 3 GOES HERE

   \section{Updates to ForC}

   CURRENT TABLE 4 GOES HERE

output:
  rticles::copernicus_article: default
  bookdown::pdf_book:
    base_format: rticles::copernicus_article # for using bookdown features like \@ref()
---

@@

@@

Previous versions of ForC [@refs] contained most of the information required for transfer to EFDB, and, more broadly, to inform C accounting under IPCC guidelines. 
However, modest changes to the structure and contents of ForC were needed in order to provide all information required by EFDB and to improve ForC's capacity to serve as a repository of valuable information for forest C accounting under IPCC guidelines.
To support export of data to EFDB, and to improve the overall quality of the ForC database, we added or modified ## fields (Table ##), defined `r n_new_variables` new variables, implemented enhanced quality control, manually reviewed ># records to obtain additional required information, and added # new records.

@@

@@

We added a total of `r n_new_variables` new variables to the set of named and defined variables, counting each pair of variables with units in C (ending in  `_C`) or organic matter (ending in `_OM`) as one. 
All of these variables are relevant to C accounting under IPCC, and hence to EFDB (Fig. `r fig_variable_mapping`)
This included eleven increment variables, adding to only one previously defined increment variable (aboveground biomass increment, *delta.agb*). 
These are directly related to C stocks as previously defined in ForC, with "*delta.*" added in front of the variable name.
Further, we added variables capturing the belowground component of woody mortality (*woody.mortality_root*) and the combined aboveground and belowground components of woody mortality (*woody.mortality*).
Although most of these variables currently lack records in ForC, the structure exists such that records can be populated over time.

@@

## Mapping ForC to EFDB
### Carbon cycle variables

@@

\clearpage
# Appendix A 
<!--remove number-->

```{r table_ForCfieldmapping, eval = TRUE, echo=FALSE, warning=FALSE}
library(knitr)
library(kableExtra)

@@

knitr::kable(ForC_variables_included, format = "latex",
             longtable = TRUE,
             caption = "\\textbf{Mapping of ForC fields to EFDB.} See footnotes at end of table (still need to be properly inserted). ") %>%
  row_spec(0, bold = TRUE) %>%
kableExtra::kable_styling(full_width = TRUE, latex_options = c("repeat_header", "hold position"))

@@

\clearpage
# Appendix B
<!--remove number-->

```{r table_ForCchanges, eval = TRUE, echo=FALSE, warning=FALSE, message = FALSE}
library (readr)
urlfile="https://raw.githubusercontent.com/forc-db/ForC/master/database_management_records/record_of_changes.csv"
record_of_changes <-read_csv(url(urlfile))
names(record_of_changes) <- stringr::str_replace_all(names(record_of_changes), "\\.", " ")
record_of_changes <- record_of_changes[c(1:5)] #include all fields but date of change
#record_of_changes<- record_of_changes[which(!grep("housekeeping",record_of_changes$Motivation)) , c(1:5)]  #trying to exclude rows for which Motivation = "housekeeping", but this is not working
record_of_changes[is.na(record_of_changes)]<- "" #replace NAs with blanks

library(knitr)
library(kableExtra)
knitr::kable(record_of_changes, format = "latex",
             longtable = TRUE,
             caption = "\\textbf{Table of changes to ForC fields.}") %>%
  kableExtra::kable_styling(full_width = TRUE, latex_options = c("repeat_header")) %>%
  row_spec(0, bold = TRUE)
```

>>>>

bibliography: references.bib
running:
  title: IPCC forest C accounting with ForC
  author: Williams et al.
competinginterests: |
  The authors declare no competing interests.
# OPTIONAL:
algorithms: true
# See https://publications.copernicus.org/for_authors/licence_and_copyright.html, normally used for transferring the copyright, if needed. 
# Note: additional copyright statements for affiliated software or data need to be placed in the data availability section. 
copyrightstatement: |
### The author's copyright for this publication is transferred to institution/company. 
### Note: unless stated otherwise, software and data affiliated with the manuscript are assumed to be published under the same licence as the article (currently Creative Commons 4.0)
availability:
  #  use this to add a statement when having only software code available
  #  use this to add a statement when having only data sets available
  #code: |
  #data: |
  codedata: |
    use this to add a statement when having data sets and software code available
authorcontribution: |
  (fill this in)
disclaimer: |
acknowledgements: |
  Thank you to all researchers who collected and published the data contained in ForC, and to all research assistants and collaborators who have helped to build the database.
  Funding for this study was provided by Bezos Earth Fund to The Nature Conservancy, the Institute for Global Environmental Strategies, WLS(?)
appendix: 
output:
  rticles::copernicus_article: default
  bookdown::pdf_book:
    base_format: rticles::copernicus_article # for using bookdown features like \@ref()

    header-includes:
    - \usepackage{caption}
---

@@

#Appendices
table_ForCchanges = "Appendix A"
table_ForCfieldmapping = "Appendix B"

@@

Previous versions of ForC [@refs] contained most of the information required for transfer to EFDB, and, more broadly, to inform C accounting under IPCC guidelines. 
However, modest changes to the structure and contents of ForC were needed in order to provide all information required by EFDB and to improve ForC's capacity to serve as a repository of valuable information for forest C accounting under IPCC guidelines.
To support export of data to EFDB, and to improve the overall quality of the ForC database, we added or modified ## fields (`r table_ForCchanges`), defined `r n_new_variables` new variables, implemented enhanced quality control, manually reviewed ># records to obtain additional required information, and added # new records.

@@

(`r table_ForCchanges`)

@@

We added a total of `r n_new_variables` new variables to the set of named and defined variables, counting each pair of variables with units in C (ending in  `_C`) or organic matter (ending in `_OM`) as one. 
All of these variables are relevant to C accounting under IPCC, and hence to EFDB (Fig. `r fig_variable_mapping`).
This included eleven increment variables, adding to only one previously defined increment variable (aboveground biomass increment, *delta.agb*). 
These are directly related to C stocks as previously defined in ForC, with "*delta.*" added in front of the variable name.
Further, we added variables capturing the belowground component of woody mortality (*woody.mortality_root*) and the combined aboveground and belowground components of woody mortality (*woody.mortality*).
Although most of these variables currently lack records in ForC, the structure exists such that records can be populated over time.

@@

## Mapping ForC to EFDB

`r table_ForCfieldmapping`

### Carbon cycle variables

@@

\clearpage
# Appendix A. Updates to ForC {-}

\captionsetup[table]{labelformat=empty}
Table A1: **Table of changes to ForC fields.**
```{r table_ForCchanges, eval = TRUE, echo=FALSE, warning=FALSE, message = FALSE}
library (readr)
urlfile="https://raw.githubusercontent.com/forc-db/ForC/master/database_management_records/record_of_changes.csv"
record_of_changes <-read_csv(url(urlfile))
names(record_of_changes) <- stringr::str_replace_all(names(record_of_changes), "\\.", " ")
record_of_changes <- record_of_changes[c(1:5)] #include all fields but date of change
#record_of_changes<- record_of_changes[which(!grep("housekeeping",record_of_changes$Motivation)) , c(1:5)]  #trying to exclude rows for which Motivation = "housekeeping", but this is not working
record_of_changes[is.na(record_of_changes)]<- "" #replace NAs with blanks

library(knitr)
library(kableExtra)
knitr::kable(record_of_changes, format = "latex",
             longtable = TRUE) %>%
  kableExtra::kable_styling(full_width = TRUE, latex_options = c("repeat_header")) %>%
  row_spec(0, bold = TRUE)
```

\clearpage
# Appendix B. Mapping ForC to EFDB {-}

Table B1: **Mapping of ForC fields to EFDB.** See footnotes at end of table (STILL NEED TO BE PROPERLY INSERTED).
```{r table_ForCfieldmapping, eval = TRUE, echo=FALSE, warning=FALSE}
library(knitr)
library(kableExtra)

@@

knitr::kable(ForC_variables_included, format = "latex",
             longtable = TRUE) %>%
  row_spec(0, bold = TRUE) %>%
kableExtra::kable_styling(full_width = TRUE, latex_options = c("repeat_header", "hold position"))

@@

### 2022-04-16 doc/manuscript/ESSDmanuscript.tex
appendix arrangement

Previous versions of ForC \citep{refs} contained most of the information
required for transfer to EFDB, and, more broadly, to inform C accounting
under IPCC guidelines. However, modest changes to the structure and
contents of ForC were needed in order to provide all information
required by EFDB and to improve ForC's capacity to serve as a repository
of valuable information for forest C accounting under IPCC guidelines.
To support export of data to EFDB, and to improve the overall quality of
the ForC database, we added or modified \#\# fields (Table \#\#),
defined 15 new variables, implemented enhanced quality control, manually
reviewed \textgreater\# records to obtain additional required
information, and added \# new records.

@@

@@

We added a total of 15 new variables to the set of named and defined
variables, counting each pair of variables with units in C (ending in
\texttt{\_C}) or organic matter (ending in \texttt{\_OM}) as one. All of
these variables are relevant to C accounting under IPCC, and hence to
EFDB (Fig. 1) This included eleven increment variables, adding to only
one previously defined increment variable (aboveground biomass
increment, \emph{delta.agb}). These are directly related to C stocks as
previously defined in ForC, with ``\emph{delta.}'' added in front of the
variable name. Further, we added variables capturing the belowground
component of woody mortality (\emph{woody.mortality\_root}) and the
combined aboveground and belowground components of woody mortality
(\emph{woody.mortality}). Although most of these variables currently
lack records in ForC, the structure exists such that records can be
populated over time.

@@

@@

\section{Appendix A}

@@

\caption{\label{tab:table_ForCfieldmapping}\textbf{Mapping of ForC fields to EFDB.} See footnotes at end of table (still need to be properly inserted). }\\

@@

\caption[]{\textbf{Mapping of ForC fields to EFDB.} See footnotes at end of table (still need to be properly inserted).  \textit{(continued)}}\\

@@

\section{Appendix B}

\begin{longtabu} to \linewidth {>{\raggedright}X>{\raggedright}X>{\raggedright}X>{\raggedright}X>{\raggedright}X}
\caption{\label{tab:table_ForCchanges}\textbf{Table of changes to ForC fields.}}\\
\hline
\textbf{Table} & \textbf{Column} & \textbf{Description} & \textbf{Changes} & \textbf{Motivation}\\
\hline
\endfirsthead
\caption[]{\textbf{Table of changes to ForC fields.} \textit{(continued)}}\\
\hline
\textbf{Table} & \textbf{Column} & \textbf{Description} & \textbf{Changes} & \textbf{Motivation}\\
\hline
\endhead
Sites & coordinates.precision & Precision of geographic coordinates, as reported by source or estimated from maps. & field added & allow identification of records with poor coordinate precision\\
\hline
Measurements & data.location.within.source & Location of data within the source listed in citation.ID. & field added & facilitate review, ensure traceability\\
\hline
 & sd, se, lower95\%CI, upper 95\%CI & Standard deviation, standard error, and lower and upper 95 percent confidence intvervals, respectively. & replaces `stat` and `stat.name` & cleaner format; ability to handle assymetrical 95 percent confidence intervals\\
\hline
 & mean.in.original.units, original.units & mean value and units presented in original publication & fields added & provide IPCC with original units, reduce errors/improve reproducibility\\
\hline
 & C.conversion.factor & Assumed/ measured C content of organic matter used to convert organic matter to C. & field added & track units conversion, allow back-calculation of OM if conversion factor deemed inappropriate\\
\hline
PFT & description & Definition of the pftcode at the community level. Differs from individual level in that properly describes mixed plant functional types. & field added & \\
\hline
 & description.individual & Definition of the pftcode at the individual plant level. & field name change (previously `description`) & \\
\hline
Citations & citation.citation & Full citation. Most of these records are automatically generated in R based upon DOI lookup. & field added & field required by IPCC\\
\hline
 & citation.language & Language of original publication, automatically generated based on the title and abstract, with some manual entries and corrections. & field added & field required by IPCC\\
\hline
 & citation.url & URL of original publication, generally retrieved automatically via URL lookup. & field added & field required by IPCC\\
\hline
 & citation.abstract & Abstract, generally retrieved automatically via DOI lookup. & field added & field required by IPCC\\
\hline
 & source.type & citation source type & field added & field required by IPCC\\
\hline
 & pdf.in.repository & Indicates whether pdf of original study has been retrieved and saved in ForC's reference repository & field added & housekeeping\\
\hline
 & EFDB.ready & Indicates whether data have been checked for export to EFDB. & field added & housekeeping\\
\hline
\end{longtabu}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\appendix
\section{Mapping ForC to EFDB}

CURRENT TABLE 3 GOES HERE

\section{Updates to ForC}

CURRENT TABLE 4 GOES HERE
\noappendix

>>>>

Previous versions of ForC \citep{refs} contained most of the information
required for transfer to EFDB, and, more broadly, to inform C accounting
under IPCC guidelines. However, modest changes to the structure and
contents of ForC were needed in order to provide all information
required by EFDB and to improve ForC's capacity to serve as a repository
of valuable information for forest C accounting under IPCC guidelines.
To support export of data to EFDB, and to improve the overall quality of
the ForC database, we added or modified \#\# fields (Appendix A),
defined 15 new variables, implemented enhanced quality control, manually
reviewed \textgreater\# records to obtain additional required
information, and added \# new records.

@@

(Appendix A)

@@

We added a total of 15 new variables to the set of named and defined
variables, counting each pair of variables with units in C (ending in
\texttt{\_C}) or organic matter (ending in \texttt{\_OM}) as one. All of
these variables are relevant to C accounting under IPCC, and hence to
EFDB (Fig. 1). This included eleven increment variables, adding to only
one previously defined increment variable (aboveground biomass
increment, \emph{delta.agb}). These are directly related to C stocks as
previously defined in ForC, with ``\emph{delta.}'' added in front of the
variable name. Further, we added variables capturing the belowground
component of woody mortality (\emph{woody.mortality\_root}) and the
combined aboveground and belowground components of woody mortality
(\emph{woody.mortality}). Although most of these variables currently
lack records in ForC, the structure exists such that records can be
populated over time.

@@

Appendix B

@@

\section*{Appendix A. Updates to ForC}
\addcontentsline{toc}{section}{Appendix A. Updates to ForC}

\captionsetup[table]{labelformat=empty}

Table A1: \textbf{Table of changes to ForC fields.}

\begin{longtabu} to \linewidth {>{\raggedright}X>{\raggedright}X>{\raggedright}X>{\raggedright}X>{\raggedright}X}
\hline
\textbf{Table} & \textbf{Column} & \textbf{Description} & \textbf{Changes} & \textbf{Motivation}\\
\hline
\endfirsthead
\multicolumn{5}{@{}l}{\textit{(continued)}}\\
\hline
\textbf{Table} & \textbf{Column} & \textbf{Description} & \textbf{Changes} & \textbf{Motivation}\\
\hline
\endhead
Sites & coordinates.precision & Precision of geographic coordinates, as reported by source or estimated from maps. & field added & allow identification of records with poor coordinate precision\\
\hline
Measurements & data.location.within.source & Location of data within the source listed in citation.ID. & field added & facilitate review, ensure traceability\\
\hline
 & sd, se, lower95\%CI, upper 95\%CI & Standard deviation, standard error, and lower and upper 95 percent confidence intvervals, respectively. & replaces `stat` and `stat.name` & cleaner format; ability to handle assymetrical 95 percent confidence intervals\\
\hline
 & mean.in.original.units, original.units & mean value and units presented in original publication & fields added & provide IPCC with original units, reduce errors/improve reproducibility\\
\hline
 & C.conversion.factor & Assumed/ measured C content of organic matter used to convert organic matter to C. & field added & track units conversion, allow back-calculation of OM if conversion factor deemed inappropriate\\
\hline
PFT & description & Definition of the pftcode at the community level. Differs from individual level in that properly describes mixed plant functional types. & field added & \\
\hline
 & description.individual & Definition of the pftcode at the individual plant level. & field name change (previously `description`) & \\
\hline
Citations & citation.citation & Full citation. Most of these records are automatically generated in R based upon DOI lookup. & field added & field required by IPCC\\
\hline
 & citation.language & Language of original publication, automatically generated based on the title and abstract, with some manual entries and corrections. & field added & field required by IPCC\\
\hline
 & citation.url & URL of original publication, generally retrieved automatically via URL lookup. & field added & field required by IPCC\\
\hline
 & citation.abstract & Abstract, generally retrieved automatically via DOI lookup. & field added & field required by IPCC\\
\hline
 & source.type & citation source type & field added & field required by IPCC\\
\hline
 & pdf.in.repository & Indicates whether pdf of original study has been retrieved and saved in ForC's reference repository & field added & housekeeping\\
\hline
 & EFDB.ready & Indicates whether data have been checked for export to EFDB. & field added & housekeeping\\
\hline
\end{longtabu}

\section*{Appendix B. Mapping ForC to EFDB}
\addcontentsline{toc}{section}{Appendix B. Mapping ForC to EFDB}

Table B1: \textbf{Mapping of ForC fields to EFDB.} See footnotes at end
of table (STILL NEED TO BE PROPERLY INSERTED).

@@

@@

\multicolumn{5}{@{}l}{\textit{(continued)}}\\

@@

@@
